---
title: "Practical Machine Learning Project Report"
author: "Punyashree PB"
output: html_document
---

###PROBLEM STATEMENT:
The goal of this project is to predict or classify new exercises using a subset of predictiors from the data (Predicting the `classe` vairalbe), using data recorded from four types of body sensors during body building exercises. After applying various data cleanup and preprossesing techniques, a training model was built to predict 20 observations from a test dataset. In this document, explained the methods and results obtained.

###DATA:
The training data for this project are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

###METHODOLOGY
####Loading Required Packages

```{r, message=FALSE}
library(rpart)
library(caret)
library(randomForest)
```

####Loading Data:
```{r}
train <- read.csv("pml-training.csv", na.strings=c("", "NA", "NULL"))
test <- read.csv("pml-testing.csv", na.strings=c("", "NA", "NULL"))
```

####Data Pre-processing
Lets remove the columns containing _NA's_ from the data
```{r}
train <-train[,colSums(is.na(train)) == 0]
test <-test[,colSums(is.na(test)) == 0]
```

Lets remove the unwanted columns fromt he data
```{r}
train   <-train[,-c(1:7)]
test <-test[,-c(1:7)]
```

####Data Partition
Lets partition the train data in to training and validation set.
```{r}
trainset <- createDataPartition(train$classe, p = 0.8, list = FALSE)
Training <- train[trainset, ]
Validation <- train[-trainset, ]
```

####Histogram Plot
Lets draw a simple histogram plot for the prediction variable.
```{r hist_plot, fig.height=4, fig.width=6}
plot(Training$classe, col="gray", 
     main="Histogram of Predicting variable(classe) in Training set", 
     xlab="classe levels", ylab="Frequency")
```

####Building Model
Lets use reandom forest for building the model.
```{r}
rfMod <- randomForest(classe ~. , data=Training, method="class")
rfMod
```

####Cross Validation.
Lets cross validate using the Validation set
```{r}
rfPred <- predict(rfMod, Validation, type = "class")
confusionMatrix(rfPred, Validation$classe)
```

The Cross validation accuracy is **99.5%** so the out of sample error is **0.5%**, which confirms our model has performed good.

####Test Prediction
Lets predict the test set using the our model _rfMod_
```{r}
testPred <- predict(rfMod, test, type="class")
testPred
```

Lets save the output as txt files as mentioned in the submission instructions using the following code, which was given as the submission instructions.

```{r}
answers <- as.character(testPred)

pml_write_files = function(x){
    n = length(x)
    for(i in 1:n){
        filename = paste0("problem_id_",i,".txt")
        write.table(x[i], file=filename, quote=FALSE, row.names=FALSE, 
                    col.names=FALSE)
    }
}

pml_write_files(answers)
```

###CONCLUSION
1. Used Random Forest to create the model. 
2. Cross Validation is done and the out of sample error is found out to be **0.5%** which is pretty good model.